\documentclass {report}
\usepackage{float}
\usepackage{longtable}
\usepackage{color}
\usepackage{appendix}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true, % make the links colored
    linkcolor=black, % color TOC links in black
    urlcolor=red, % color URLs in red
    linktoc=all % 'all' will create links for everything in the TOC
}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{csquotes}
\usepackage{fancyvrb}


\title {
	{sidBison: A Stepwise Interactive Debugger for Bison\\[0.4in]}
	{ Tufts University\\[0.2in]}
	{\large Advisor: Prof. Kathleen Fisher\\}
	}
\author {Siddhartha Prasad}
\date{21st March 2016}


\begin {document}

\maketitle

\input{Abstract.tex}

\pagebreak
\tableofcontents
\pagebreak


\thispagestyle{plain}
\begin{center}
    \textbf{A Quick Note}
\end{center}

\noindent

This thesis is aimed at audiences who have previously encountered formal grammars and/or concrete syntax trees.
A ppendix A contains a quick introduction to formal and context-free grammars.\\

Familiarity with parser-generators like \verb|Yacc| or \verb|Bison| is not required, but will allow for deeper understanding of the problem. I recommend reading the Bison user manual at \verb|http://www.gnu.org/software/bison/manual/|.

\pagebreak

\input{Motivation.tex}

\input{Goals.tex}

\input{TechnicalApproach.tex}

\input{Complexity.tex}

\input{Evaluation.tex}


\input{AdditionalWork.tex}

\input{app.tex}



\begin{thebibliography}{}
\bibitem{Appel} Appel, Andrew W. Modern Compiler Implementation in ML. Cambridge: Cambridge UP, 1998. Print.

\bibitem {Donnelly and Stallman} Donnelly, Charles, and Richard Stallman. "Bison The YACC-compatible Parser Generator." Bison, The YACC-compatible Parser Generator. N.p., Nov. 1995. Web. 26 Sept. 2015. <http://dinosaur.compilertools.net/bison/index.html>.

\bibitem{Understanding Bison} Popuri, Satya K. "Understanding C Parsers Generated by GNU Bison." Understanding Parsers Generated by GNU Bison. N.p., n.d. Web. 17 Feb. 2016.

\bibitem{iBison} Popuri, Satya K. "Interactive Mode Bison." Interactive Parsing with GNU Bison. University of Illinois, Chicago, n.d. Web. 17 Feb. 2016. \verb|https://www.cs.uic.edu/~spopuri/ibison.html|

\bibitem{Bidwell} Bidwell, Daniel R. "History of Parsing Methods." History of Parsing Methods. N.p., n.d. Web. 17 Feb. 2016. \verb|https://www.andrews.edu/~bidwell/456/history.html|

\bibitem{Bison Algorithm} "The Bison Parser Algorithm." Bison 1.25. N.p., n.d. Web. 17 Feb. 2016.\verb|http://dinosaur.compilertools.net/bison/bison_8.html|

\bibitem {Bray Tim} Bray, Tim. "The JavaScript Object Notation (JSON) Data Interchange Format." The JavaScript Object Notation (JSON) Data Interchange Format. N.p., n.d. Web. 07 Oct. 2015. http://rfc7159.net/rfc7159.

\bibitem{LRWiki} "LR Parser." Wikipedia. Wikimedia Foundation, n.d. Web. 28 Mar. 2015.

\bibitem{BNF} Lars Garshol. "BNF and EBNF: What Are They and How Do They Work?" BNF and EBNF: What Are They and How Do They Work? N.p., 16 June 1998. Web. 18 Feb. 2016.

\bibitem{GNU} "Bison - GNU Project - Free Software Foundation." Bison - GNU Project - Free Software Foundation. N.p., n.d. Web. 19 Feb. 2016. https://www.gnu.org/software/bison/

\bibitem{Yacc is dead} Might, Matthew, and David Darais. "Yacc Is Dead." (n.d.): n. pag. University Of Utah, Salt Lake City, Utah, Usa,. Web.

\bibitem{Complexity GLR} Johnson, Mark. "Chapter 3: The Computational Complexity of GLR Parsing." Generalized LR Parsing. Ed. Masaru Tomita. Boston: Kluwer Academic, 1991. N. pag. Print. 

\bibitem{LinearBison} \verb|http://www.gnu.org/software/bison/manual/html_node/Simple-GLR-Parsers.html|

\end{thebibliography}

\end{document}



